---
title: "Conditional Flow"
author: "Rochelle Terman"
date: "Fall 2020"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r message=F}
library(tidyverse)
library(gapminder)
library(knitr)
```

## Conditional Flow 

Sometimes you only want to execute code if a certain condition is met. To do that, we use an **if-else statement.** It looks like this:

```{r eval = F}
if (condition) {
  # code executed when condition is TRUE
} else {
  # code executed when condition is FALSE
}
```

`condition` is a statement that must always evaluate to either `TRUE` or `FALSE`. This is similar to `filter()`, except `condition` can only be a single value (i.e. a vector of length 1), whereas `filter()` works for entire vectors (or columns).

Let's look at a simple example:

```{r}
age = 84
if (age > 60) {
    print("OK Boomer")
} else {
    print("But you don't look like a professor!")
}
```

We refer to the first `print` command as the first *branch*. 

Let's change the `age` variable to execute the second branch:

```{r}
age = 20
if (age > 60) {
    print("OK Boomer")
} else {
    print("But you don't look like a professor!")
}
```

### Multiple Conditions

You can chain conditional statements together:

```{r eval = F}
if (this) {
  # do that
} else if (that) {
  # do something else
} else {
  # do something completely different
}
```

### Complex Statements

We can generate more complex conditional statements with boolean operators like `&` and `|`:

```{r}
age = 45 

if (age > 60) {
    print("OK Boomer")
} else if (age < 60 & age > 40) {
    print("How's the midlife crisis?")
} else {
    print("But you don't look like a professor!")
}
```

### Code Style

Both `if` and `function` should (almost) always be followed by squiggly brackets (`{}`), and the contents should be indented. This makes it easier to see the hierarchy in your code by skimming the left-hand margin.

An opening curly brace should never go on its own line and should always be followed by a new line. A closing curly brace should always go on its own line, unless it’s followed by else. Always indent the code inside curly braces.

```{r eval=F}
# Bad
if (y < 0 && debug)
message("Y is negative")

if (y == 0) {
  log(x)
} 
else {
  y ^ x
}

# Good
if (y < 0 && debug) {
  message("Y is negative")
}

if (y == 0) {
  log(x)
} else {
  y ^ x
}
```

### `if` vs. `if_else`

Because if-else conditional statements like the ones outlined above must always resolve to a single `TRUE` or `FALSE`, they cannot be used for vector operations. Vector operations are where you make multiple comparisons simultaneously for each value stored inside a vector. 

Consider the `gapminder` data and imagine you wanted to create a new column identifying whether or not a country-year observation has a life expectancy of at least 35.

```{r}
gap <- gapminder
kable(head(gap))
```

This sounds like a classic if-else operation. For each observation, if `lifeExp` is greater than or equal to `35`, then the value in the new column should be `1`. Otherwise, it should be `0`. But what happens if we try to implement this using an if-else operation like above?

```{r}
gap_if <- gap %>%
   mutate(life.35 = if(lifeExp >= 35){
     1
   } else {
     0
   })

head(gap_if)
```

This did not work correctly. Because `if()` can only handle a single `TRUE/FALSE` value, it only checked the first row of the data frame. That row contained `28.801`, so it generated a vector of length 1704 with each value being `0`.

Because we in fact want to make this if-else comparison 1704 times, we should instead use **`if_else()`**. This **vectorizes** the if-else comparison and makes a separate comparison for each row of the data frame. This allows us to correctly generate this new column.

```{r}
gap_ifelse <- gap %>%
  mutate(life.35 = if_else(lifeExp >= 35, 1, 0))

kable(head(gap_ifelse))
```


