---
title: "Functions"
author: "Rochelle Terman"
date: "Fall 2020"
output: html_document
---

```{r message=F}
library(knitr)
library(tidyverse)
library(gapminder)
```

## Functions

Functions are the basic building blocks of programs. Think of them as "mini-scripts" or "tiny commands." We've already used dozens of functions created by others (e.g. `filter()`, `mean()`.) 

This lesson teaches you how to write you own functions, and why you would want to do so.  The details are pretty simple, but this is one of those ideas where it's good to get lots of practice!

### Why Write Functions? 

Functions allow you to automate common tasks in a more powerful and general way than copy-and-pasting. For example, take a look at the following code:

```{r eval = F}
gap <- gapminder

kable(head(gapminder))

gap_norm <- gap %>%
  mutate(pop_norm = (pop - min(pop)) / (max(pop) - min (pop)),
         gdp_norm = (gdpPercap - min(gdpPercap)) / (max(gdpPercap) - min (gdpPercap)),
         life_norm = (lifeExp - min(lifeExp) / (max(pop)) - min (lifeExp)))

summary(gap_norm$pop_norm)
```

You might be able to puzzle out that this rescales each numeric column to have a range from 0 to 1. But did you spot the mistakes? I made two errors when copying-and-pasting the code for `lifeExp`.

Functions have a number of advantages over this "copy-and-paste" approach:

- **They are easy to reuse.** If you need to change things, you only have to update code in one place, instead of many.

- **They are self-documenting.** Functions name pieces of code the way variables name strings and numbers. Give your function a good name and you will easily remember the function and its purpose.

- **They are easier to debug.** There are fewer chances to make mistakes because the code only exists in one location (i.e. updating a variable name in one place, but not in another).

### Anatomy of a Function

Functions have three key components:

1. A **name**. This should be informative and describe what the function does.

2. The **arguments**, or list of inputs, to the function. They go inside the parentheses in `function()`.

3. The **body**. This is the block of code within `{}` that immediately follows `function(...)`, and is the code that you develop to perform the action described in the name using the arguments you provide.

```{r eval = F}
my_function <- function(x, y){
  # do
  # something
  # here
  return(result)
}
```

In this example, `my_function` is the **name** of the function, `x` and `y` are the **arguments**, and the stuff inside the `{}` is the **body**.

### Writing a Function

Let's re-write the scaling code above as a function. To write a function you need to first analyze the code. How many inputs does it have?

```{r}
# the corrected code
gap_norm <- gap %>%
  mutate(pop_norm = (pop - min(pop)) / (max(pop) - min (pop)),
         gdp_norm = (gdpPercap - min(gdpPercap)) / (max(gdpPercap) - min (gdpPercap)),
         life_norm = (lifeExp - min(lifeExp)) / (max(lifeExp) - min (lifeExp)))

# focus on the line:
# pop_norm = (pop - min(pop)) / (max(pop) - min (pop))
```

This code only has one input: `gap$pop`. To make the inputs more clear, it's a good idea to rewrite the code using temporary variables with general names. Here this code only requires a single numeric vector, which I’ll call `x`:

```{r}
x <- gap$pop

(x - min(x)) / (max(x) - min(x))
```

There is still some duplication in this code. We're calulating some version of the range three times. Pulling out intermediate calculations into named variables is a good practice because it becomes more clear what the code is doing. 

```{r}
x <- gap$pop

rng <- range(x)
(x - rng[1]) / (rng[2] - rng[1])
```

Now that I’ve simplified the code, and checked that it still works, I can turn it into a function:

```{r}
rescale01 <- function(x) {
  rng <- range(x)
  scales <- (x - rng[1]) / (rng[2] - rng[1])
  return(scales)
}
```

Note the overall process: I only made the function after I’d figured out how to make it work with a simple input. It’s easier to start with working code and turn it into a function; it’s harder to create a function and then try to make it work.

At this point it’s a good idea to check your function with a few different inputs:

```{r}
rescale01(c(-10, 0, 10))

rescale01(c(1, 2, 3, 5))
```

### Using a Function

Two important points about using (or *calling**) functions:

1. Notice that when we **call** a function, we're passing a value into it that is assigned to the parameter we defined when writing the function. In this case, the parameter `x` is automatically assigned to `c(-10, 0, 10)`.

2. When using functions, by default the returned object is merely printed to the screen. If you want it saved, you need to assign it to an object.

Let's see if we can simplify the original example with our brand new function:

```{r eval = F}
rescale01 <- function(x) {
  rng <- range(x)
  scales <- (x - rng[1]) / (rng[2] - rng[1])
  return(scales)
}

gap_norm <- gap %>%
  mutate(pop_norm = rescale01(pop),
         gdp_norm = rescale01(gdpPercap),
         life_norm = rescale01(lifeExp))

```

Compared to the original, this code is easier to understand and we’ve eliminated one class of copy-and-paste errors. There is still quite a bit of duplication since we’re doing the same thing to multiple columns. We’ll learn how to eliminate that duplication in the lesson on iteration.

Another advantage of functions is that if our requirements change, we only need to make the change in one place. For example, we might discover that some of our variables include `NA` values, and `rescale01()` fails:

```{r}
rescale01(c(1, 2, NA, 3, 4, 5))
```

Because we’ve extracted the code into a function, we only need to make the fix in one place:

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = T)
  scales <- (x - rng[1]) / (rng[2] - rng[1])
  return(scales)
}

rescale01(c(1, 2, NA, 3, 4, 5))
```


### Challenges

#### Challenge 1 {-}

Write a function that calculates the sum of the squared value of two numbers. For instance, it should generate the following output:

```{r eval = F}
my_function(3, 4)
# [1] 25

```

#### Challenge 2 {-}

Write `both_na()`, a function that takes two vectors and returns the total number NAs in both vectors. 

For instance, it should generate the following output:

```{r eval = F}
vec1 <- c(NA, 4, 6, 2, NA, 5, NA)
vec2 <- c(NA, "Dec", "Apr", NA, "Jul", "Apr")

my_other_function(vec1, vec2)
# [1] 5

# hints
is.na(vec1)
sum(c(T, F))

```

#### Challenge 3 {-}

Fill in the blanks to create a function that takes a name like "Rochelle Terman" and returns that name in uppercase and reversed, like "TERMAN, ROCHELLE"

```{r eval = F}
standard_names <- function(name){
    upper_case = toupper(____) # make upper
    upper_case_vec = strsplit(_____, split = ' ')[[1]] # turn into a vector
    first_name = ______ # take first name
    last_name = _______ # take last name
    reversed_name = paste(______, _______, sep = ", ") # reverse and separate by a comma and space
    return(reversed_name)
}

```

#### Challenge 4 {-}

Look at the following function:

```{r}
print_date <- function(year, month, day){
    joined = paste(as.character(year), as.character(month), as.character(day), sep = "/")
    return(joined)
}
```

Why do the two lines of code below return different values? 

```{r eval = F}
print_date(day=1, month=2, year=2003)
print_date(1, 2, 2003)
```

