---
title: "Day 2: Challenge Solutions"
subtitle: "ESS 3I: Webscraping and Data Management in R"
author: "Rochelle Terman"
date: "Fall 2019"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(knitr)
library(gapminder)
```

## Strings

#### Challenge 1.

In code that doesn't use `stringr`, you'll often see `paste()` and `paste0()`. What's the difference between the two functions? What `stringr` function are they equivalent to? How do the functions differ in their handling of  `NA`?

- `paste()` adds a space separater by default; `paste0()` does not. 
- The `stringr` equivalent to `paste()` is `str_c()`, as in `str_c("hey", "you", sep = " ")`. 
- `paste()` and `paste0()` will convert `NA`s to characters; `str_c()` will not.
    
#### Challenge 2. 

In your own words, describe the difference between the `sep` and `collapse` arguments to `str_c()`.

- `sep = ","` inserts a "," between vector arguments. Output is a vector.
- `collapse = ","` will combine input vectors into a single string, Output is a single string (i.e. a vector of length = 1)

```{r}
vec1 <- c("a", "b", "c")
vec2 <- c(1, 2, 3)

str_c(vec1, vec2, sep = ".")
# c("a.1" , "b.2" , "c.3")

str_c(vec1, vec2, collapse = ".")
# "a1.b2.c3"
```


#### Challenge 3.

Use `str_length()` and `str_sub()` to extract the middle character from 
a string. What will you do if the string has an even number of characters?

```{r}
# for odd cases
x <- "lemon"
str_sub(x,
        start = ceiling(str_length(x) / 2 ), 
        end = ceiling(str_length(x) / 2))

# for both odd and even cases
x <- "pies"
str_sub(x,
        start = ifelse(str_length(x) %% 2 == 0, 
                       floor(str_length(x) / 2), 
                       ceiling(str_length(x) / 2 )), 
        end = floor(str_length(x) / 2) + 1)
```


#### Challenge 4. 

What does `str_trim()` do? What's the opposite of `str_trim()`?

- `str_trim()` removes whitespace from start and end of string.
- `str_pad()` adds whitespace to star tand eng of string.

## Functions

#### Challenge 

Write a function that calculates the sum of the squared value of two numbers. For instance, it should generate the following output:

```{r eval = F}
# SOLUTION
my_function <- function(donut, kitten) {
  sum <- sum(donut^2,  kitten^2)
  return(sum)
}

my_function(3, 4)
my_function(1,2)
```

#### Challenge 2 

Write `both_na()`, a function that takes two vectors and returns the total number NAs in both vectors. 

For instance, it should generate the following output:

```{r eval = F}
# SOLUTION
both_na <- function(x, y) {
  sum(is.na(x), is.na(y))
}

both_na(
  c(NA, NA, 1, 2, NA, NA, 1),
  c(NA, 1, NA, 2, NA, NA, 1)
)
```

#### Challenge 3 

Fill in the blanks to create a function that takes a name like "Rochelle Terman" and returns that name in uppercase and reversed, like "TERMAN, ROCHELLE"

```{r eval = F}
# SOLUTION
standard_names <- function(name){
    upper_case = toupper(name) # make upper
    upper_case_vec = strsplit(upper_case, split = ' ')[[1]]  # turn into a list
    first_name = upper_case_vec[1]# take first name
    last_name = upper_case_vec[2] # take last name
    reversed_name = paste(last_name, first_name, sep = ", ") # reverse and separate by a comma and space
    return(reversed_name)
}
standard_names('Rochelle Terman')
```

#### Challenge 4

Look at the following function:

```{r}
print_date <- function(year, month, day){
    joined = paste(as.character(year), as.character(month), as.character(day), sep = "/")
    return(joined)
}
```

Why do the two lines of code below return different values? 

```{r eval = F}
print_date(day=1, month=2, year=2003)
print_date(1, 2, 2003)
```

Line 81 uses keyword arguments; line 82 uses positional arguments. So `year` is assigned to 1, `month` to 2, and `day` to 2003.

## Iteration

### Challenges

Write code that uses one of the map functions to:

#### Challenge 1

Calculates the arithmetic mean for every column in `mtcars`.

```{r}
map(mtcars, mean)
```

#### Challenge 2

Calculates the number of unique values in each column of `iris`. 

```{r}
iris %>% 
  map(unique) %>%
  map_dbl(length)
```

#### Challenge 3

Generate 10 random normals for each of $\mu = -10$, $0$, $10$, and $100$.

```{r}
mus <- c(-10, 0, 10, 100)
map(mus, ~rnorm(n = 10, mean = .))
```