---
title: "01_Data_Wrang Solutions"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
gap <- gapminder
```

## 1_Data-Basics Challenges

#### Challenge 1. {-}

Read the `polity` dataset.

```{r}
polity <- read.csv("Data/polity_sub.csv", stringsAsFactors = F)
```

#### Challenge 2. {-}

Report the number and names of each variable in the dataset.

```{r}
dim(polity)
names(polity)
```

#### Challenge 3. {-}

What is the mean `polity2` score in the dataset?

```{r}
mean(polity$polity2, na.rm = TRUE)
```

#### Challenge 4. {-}

What is the range of the `polity2` variable?

```{r}
range(polity$polity2, na.rm = T)
```

#### Challenge 5. {-}

How many unique countries are in the dataset?

```{r}
length(unique(polity$country))
```

## 2_dplyr Challenges

#### Challenge 1. {-}

Use `dplyr` to create a data frame containing the median `lifeExp` for each continent.

```{r}
gap %>%
  group_by(continent) %>%
  mutate(median = median(lifeExp))
```

#### Challenge 2. {-}

Use `dplyr` to add a column to the gapminder dataset that contains the total population of the continent of each observation in a given year. For example, if the first observation is Afghanistan in 1952, the new column would contain the population of Asia in 1952. 

```{r warning=F}
gap %>%
  group_by(continent, year) %>%
  mutate(continent_pop = sum(as.numeric(pop)))
```

#### Challenge 3. {-}

Use `dplyr` to: (a) add a column called `gdpPercap_diff` that contains the difference between the observation's `gdpPercap` and the mean `gdpPercap` of the continent in that year, (b) arrange the dataframe by the column you just created, in descending order (so that the relatively richest country-years are listed first).

**hint**: You might have to `ungroup()` before you `arrange()`.

```{r}
gap <- gap %>%
  group_by(continent, year) %>%
  mutate(mean_gdpPercap = mean(gdpPercap)) %>%
  mutate(gdpPercap_diff = gdpPercap - mean_gdpPercap) %>%
  ungroup() %>%
  arrange(desc(gdpPercap_diff), year)
head(gap)

# Top observation should be Kuwait in 1957, followed by Kuwait in 1952, 1972, etc
```

## 3_tidyr Challenges

#### Challenge 1. {-}

Subset the results from Challenge #3 (of the previous chapter) to select only the `country`, `year`, and `gdpPercap_diff` columns. Use `tidyr` to put it in wide format so that countries are rows and years are columns.


```{r}
# Solution from 2, Challenge 3:
gap <- gap %>%
  group_by(continent, year) %>%
  mutate(mean_gdpPercap = mean(gdpPercap)) %>%
  mutate(gdpPercap_diff = gdpPercap - mean_gdpPercap) %>%
  ungroup() %>%
  arrange(desc(gdpPercap_diff), year)

# Spreading into wide format:
gap <- gap %>%
  select(country, year, gdpPercap_diff) %>%
  spread(year, gdpPercap_diff)
```

#### Challenge 2. {-}
Now turn the dataframe above back into the long format with three columns: `country`, `year`, and `gdpPercap_diff`.

```{r}
gap <- gap %>%
  gather(year, gdpPercap_diff, 2:13)
```

## 4_Relational-Data Challenges

```{r}
polity <- read.csv("data/polity_sub.csv", stringsAsFactors = F)
ciri <- read.csv("data/ciri_sub.csv", stringsAsFactors = F)

gap1 <- gapminder %>%
  left_join(polity)
```

#### Challenge 1. {-}

Merge the polityVI and CIRI datasets, keeping all observations in polityVI. Save this merged dataframe as `p1`. How many observations does `p1` have? Why?

```{r}
p1 <- polity %>%
  left_join(ciri, by = c("country" = "CTRY", "year" = "YEAR"))
```

`p1` has 16,727 observations, because `polity` has 16,727 observations, and `left_join()` keeps all observations in `x`.

#### Challenge 2. {-}

Merge the `gap1` dataset we created above with the `ciri` dataset, this time keeping all observations in `ciri`. Save this as `gap2`. How many observations does it have? What is the major problem with merging the datasets this way?

```{r}
gap2 <- ciri %>%
  left_join(gap1, by = c("CTRY" = "country", "YEAR" = "year"))
```

`gap2` has 6,262 observations. Because the ciri dataframe has many more observations than `gap1`, merging in this direction introduces a lot of `NA`s into the data.