---
title: "Relational Data"
author: "Rochelle Terman"
date: "Fall 2020"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
rm(list = ls())

library(knitr)

set.seed(1014)
options(digits = 3)
options(dplyr.print_min = 6, dplyr.print_max = 6)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)
```

```{r message=FALSE} 
library(tidyverse)
library(gapminder)
```

# Relational Data

It is rare that data analysis involves only a single table of data. Typically, you have many tables of data, and you must combine them to answer the questions that you are interested in. Collectively, multiple tables of data are called __relational data__ because it is the relations, not just the individual datasets, that are important.

Note that when we say relational database here, we are referring to how the data are structured, not to the use of any fancy software. 

## Why Relational Data

As social scientists, we are often working with data across different levels of analysis. The main principle of relational data is that each table is structured around the same observational unit. 

Why is this important? Check out the following data. 

```{r}
messy <- data.frame(
  county = c(36037, 36038, 36039, 36040, NA , 37001, 37002, 37003),
  state = c('NY', 'NY', 'NY', NA, NA, 'VA', 'VA', 'VA'),
  cnty_pop = c(3817735, 422999, 324920, 143432, NA, 3228290, 449499, 383888),
  state_pop = c(43320903, 43320903, NA, 43320903, 43320903, 7173000, 7173000, 7173000),
  region = c(1, 1, 1, 1, 1, 3, 3, 4)
)

kable(messy)
```

What a mess! How can the population of the state of New York be 43 million for one county but "missing" for another? If this is a dataset of counties, what does it mean when the "county" field is missing? If region is something like Census region, how can two counties in the same state be in different regions? And why is it that all the counties whose codes start with 36 are in New York except for one, where the state is unknown?

If we follow the principles of relational data, each type of observational unit should form a table:

* `counties` contains data on counties.
* `states` contains data on states.

So our data should look like this:

```{r}
counties <- data.frame(
  county = c(36037, 36038, 36039, 36040, 37001, 37002, 37003),
  state = c('NY', 'NY', 'NY', 'NY', 'VA', 'VA', 'VA'),
  county_pop = c(3817735, 422999, 324920, 143432, 3228290, 449499, 383888), stringsAsFactors = F
)
kable(counties)

states <- data.frame(
  state = c("NY", "VA"),
  state_pop = c(43320903, 7173000),
  region = c(1, 3), stringsAsFactors = F
)
kable(states)
```

County population is a property of a county, so it lives in the county table. State population is a property of a state, so it cannot live in the county table. If we had panel data on counties, we would need separate tables for things that vary at the county level (like state) and things that vary at the county-year level (like population).

Now the ambiguity is gone. Every county has a population and a state. Every state has a population and a region. There are no missing states, no missing counties, and no conflicting definitions. The database is self-documenting. 

## Keys

The variables used to connect each pair of tables are called **keys**. A key is a variable (or set of variables) that uniquely identifies an observation; it can also be called a *unique identifier*. 

* Keys are complete. They never take on missing values.
* Keys are unique. They are never duplicated across rows of a table.

In simple cases, a single variable is sufficient to identify an observation. In the example above, each county is identified with **county** (a numeric identifier); each state is identified with **state** (a two-letter string).

There are two types of keys:

* A **primary key** uniquely identifies an observation in its own table. For example, `counties$county` is a primary key because it uniquely identifies each county in the counties table.

* A **foreign key** uniquely identifies an observation in another table. For example, `counties$state` is a foreign key because it appears in the counties table where it matches each county to a unique state.

A primary key and the corresponding foreign key in another table form a **relation**.

Sometimes a table does not have an explicit primary key: each row is an observation, but no combination of variables reliably identifies it. If a table lacks a primary key, it is useful to add one with `mutate()` and `row_number()`. This is called a __surrogate key__.


## Joins

Data stored in the form we have outlined above is considered _normalized_. In general, we should try to keep data normalized as far into the code pipeline as we can. Storing normalized data means your data will be easier to understand and it will be harder to make costly mistakes.

At some point, however, we are going to have to merge (or __join__) the tables together to produce and analyze a single dataframe. 

Let's say we wanted to merge tables `x` and `y`. __join__ allows us to combine variables from the two tables. It first matches observations by their keys, then copies across variables from one table to the other.

There are five join options:

1. An __inner join__ keeps observations that appear in both tables.
2. A __left join__ keeps all observations in `x`.
3. A __right join__ keeps all observations in `y`.
4. A __full join__ keeps all observations in `x` and all observations in `y`.
5. An __anti join__ keeps all observations in `x` that do not have a match in `y`.

The most commonly used join is the `left_join()`: you use this whenever you look up additional data from another table, because it preserves the original observations even when there is not a match. For example, a `left_join()` on `x` and `y` pulls in variables from `y` while preserving all the observations in `x`.

Let's say we want to combine the `countries` and `states` tables we created earlier. 

```{r}
counties_states <- counties %>%
  left_join(states, by = "state")

kable(counties_states)
```

Notice there are two new columns: `state_pop` and `region`.

The left join should be your default join: use it unless you have a strong reason to prefer one of the others.

## Defining Keys

In the example above, the two tables were joined by a single variable, and that variable has the same name in both tables. That constraint was encoded by `by = "key"`. 

You can use other values for `by` to connect the tables in other ways:

1. The default, __`by = NULL`__, uses all variables that appear in both tables, what we might call a "natural join." 

For example, let's say we wanted to add a column to the `gapminder` dataset that encodes the regime type of each country-year observation. We will get that data from the `polityIV` dataset.

```{r}
gap <- gapminder
polity <- read.csv("data/polity_sub.csv", stringsAsFactors = F)
kable(head(polity))
```

We are now ready to join the tables. The common keys between them are `country` and `year`:

```{r}
gap1 <- gapminder %>%
  left_join(polity)

kable(head(gap1))
```

2. A character vector, __`by = c("x", "y")`__. This is like a natural join, but it uses only some of the common variables.

3. A named character vector: __`by = c("a" = "b")`__. This will match variable `a` in table `x` to variable `b` in table `y`. The variables from `x` will be used in the output.

For example, let's add another variable to our `gapminder` dataset -- physical integrity rights -- from the CIRI dataset.

```{r}
ciri <- read.csv("data/ciri_sub.csv", stringsAsFactors = F)
kable(head(ciri))
```

Both datasets have country and year columns, but they are named differently.

```{r}
gap2 <- gap1 %>%
  left_join(ciri, by = c("country" = "CTRY", "year" = "YEAR"))

kable(head(gap2))
```

Notice that `PHYSINT` is `NA` in the first 6 rows because the `ciri` dataset does not contain observations for Afghanistan in these years. But since we used `left_join()`, all observations in `gapminder` were preserved.

We can see some values for `PHYSINT` if we peak at the bottom of the dataset:

```{r}
kable(tail(gap2))
```

## Duplicate Keys

So far we have assumed that the keys are unique, but that is not always the case. For example, 

```{r}
x <- data.frame(key = c(1, 2),
               val_y = c("x1", "x2"))

y <- data.frame(key = c(1, 2, 2, 1),
               val_x = c("y1", "y2", "y3", "y4"))

left_join(x, y, by = "key")
```

Notice that this can sometimes cause unintended duplicates.

## Challenges

#### Challenge 1. {-}

Merge the polityVI and CIRI datasets, keeping all observations in polityVI. Save this merged dataframe as `p1`. How many observations does `p1` have? Why?

#### Challenge 2. {-}

Merge the `gap1` dataset we created above with the `ciri` dataset, this time keeping all observations in `ciri`. Save this as `gap2`. How many observations does it have? What is the major problem with merging the datasets this way?

#### Acknowledgements {-}

This page is in part derived from the following sources:

1. [R for Data Science](https://r4ds.had.co.nz), licensed under [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](https://creativecommons.org/licenses/by-nc-nd/3.0/us/).

2. [Gentzkow, Matthew and Jesse M. Shapiro. 2014. Code and Data for the Social Sciences: A Practitioner’s Guide.](https://web.stanford.edu/~gentzkow/research/CodeAndData.pdf).
