---
title: "Lab 1"
subtitle: "3I: Webcraping & Data Management in R"
author: "Rochelle Terman"
date: "August 2020"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

# Lab 1

For this lab, we'll be working with a dataset containing U.S. news coverage of women around the world (excluding the United States). In this dataset, the unit of observation is an *article* (each row represents an individual article) with variables for:

* `publication`: NYT or Washington Post.
* `year`: Year article was published.
* `title`: Title of the article.
* `country`: Country the article is mainly about.
* `region`: Region where `country` is located.
* `ccode`: Numerical code for `country`.

We'll be merging these data with another dataset where the unit of observation is *country-year.* This dataset has the following variables:

* `year`: Year.
* `ccode`: Country numerical code.
* `country`: Country name.
* `idealpoint` UN Ideal point.
* `polity2`: Polity2 score of regime type (from Polity IV).
* `physint`: Physical Integrity Rights score (from CIRI dataset).
* `speech`: Speech Rights score (from CIRI).
* `gdp.pc.wdi`: GDP per capita (from World Bank).
* `pop.wdi`: Population (from World Bank).
* `amnesty`: Political Terror Scale using Amnesty International reports (from Political Terror Scale project).
* `cinc`: Composite Index of Military Capabilities (Correlates of War).
* `region`: Geographic region.

## 1. Loading, subsetting, summarizing

### 1.1

Load the csv found in `data/articles.csv` into R. Be sure to set `stringsAsFactors` to `FALSE`. Store the dataframe to an object called `articles` and tell me the variable names.

```{r}
# YOUR CODE HERE
```

### 1.2

How many unique countries are covered in the dataset?

```{r}
# YOUR CODE HERE
```

### 1.3

The variable `ccode` reports a numerical ID corresponding to a given country. Print the names of the country or countries without a `ccode` (i.e. those countries where the `ccode` is `NA`.)

**Hint**: The `is.na()` function will return a logical vector indicating which elements are missing. `!is.na()` returns the opposite.

```{r}
# YOUR CODE HERE
```

### 1.4

Remove all articles where the `ccode` variable is `NA`. How many observations are left with? Store the trimmed dataset in an objet called `articles_no_na`

```{r}
# YOUR CODE HERE
```

## 2. Counting Frequencies and Merging

### 2.1 

With `articles_no_na`, create a new data frame called `articles_country_year` that tells us the number of articles per `ccode` (i.e. country code), per `year`.

The final data frame `articles_country_year` should contain three columns: `ccode`, `year`, and `number_articles`.

```{r}
# YOUR CODE HERE
```

### 2.2. 

Load `data/country-year.csv`.

```{r}
# YOUR CODE HERE
```

### 2.3  

Subset `country_year` such that it has the same year range as `articles_country_year`

```{r}
# YOUR CODE HERE
```

### 2.4  

Merge (i.e. join) `articles_country_year` and `country_year` into a new dataframe called `merged`.

When you're done, `merged` should have all the rows and columns of the country_year dataset, along with a new column called `number_articles`.

```{r}
# YOUR CODE HERE
```

### 2.5 

In `merged`, replace all instances of `NA` in the `number_articles` column to 0.

**hint**: Checkout the `replace_na()` function.

```{r}
# YOUR CODE HERE
```

### 2.6  

Which country-year observation has the most number of articles? Write code that prints the year, country name, and number of articles for this observation.

**hint**: Checkout the `top_n()` function.

```{r}
# YOUR CODE HERE
```


## 3. Group-wise Operations

### 3.1 

Using the `merged` data and our **split-apply-combine** strategies, print the total number of articles about women per region.

```{r}
# YOUR CODE HERE
```

## 4. Long v. wide formats

### 4.1 

Create a piped operation on `merged` that does the following:

  1. Subsets the dataframe to select `year`, `country`, and `number_articles` columns.
  2. Filters the dataframe to select only observations in the MENA region.
  3. Spreads the dataframe so that each `country` is a column, and the cells represent `number_articles.

```{r}
# YOUR CODE HERE
```

### 4.2 

Transform the dataset you created above back into `long` format, with three variables: `year`, `country`, and `number_articles`

```{r}
# YOUR CODE HERE
```

## 5. Extra Bonus Challenge

Transform the `country_year` data into an undirected dyadic dataset. Here, the unit of observation should be the dyad-year, with five columns:

  1. `ccode_1`: Country 1 `ccode`
  2. `country_1`: Country 1 name
  3. `ccode_2`: Country 2 `ccode`
  4. `country_2`: Country 2 name
  5. `year`: Year of observation
  3. `gdp_diff`: Absolute difference of gdp between dyad. 

This should be *undirected* dyadic dataset, meaning `USA-Canada-1980` is the same as `Canada-USA-1980`, and we shouldn't have duplicate rows for the same dyad.

Try to do it all in 1 piped sequence. Then tell me the dyad-year with the greatest wealth disparity.

```{r}
# YOUR CODE HERE
```